Suitability of the local-maximum entropy approximation under the Material Point Method framework for dynamic problems. \thanks{Funding: This study was funded by Agustín de Betancourt Foundation (grant number 262390106114).}

### 2020-02-21 _region_.tex
Cambios 21-2-2020

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
% flush right qed marks, e.g. at end of proof
\smartqed

@@

% Divergence
\newcommand{\Div}[1]{
  \ensuremath{\nabla \cdot {#1}}
}
% Gradient
\newcommand{\Grad}[1]{
  \ensuremath{\nabla{#1}}
}

@@

% Integral
\newcommand{\Integral}[2]{
  \IfStrEqCase{#1}{
    {1}{\ensuremath{\int_{\Gamma}{#2}\ d\Gamma}}
    {2}{\ensuremath{\int_{\Sigma}{#2}\ d\Sigma}}
    {3}{\ensuremath{\int_{\Omega}{#2}\ d\Omega}}
  }
}

@@

@@

\section{Overview of the MPM}
\label{sec:mpm-overview}
In this section, we provide an overview of the standard explicit MPM
algorithm as presented by \cite{Sulsky1994}. The method has three main
steps: a variational recovery process to project particle data to the
nodes of the grid, a Eulerian step where balance of momentum is
compute by solving internal and external forces in the nodes, and
finally the advection of the Lagrangian particles. We will first give
a introduction to the governing equations and then follow by an
algorithmic description of the process.

\subsection{Derivation of the MPM}
\label{sec:derivation-mpm}
In the MPM approach a continuum is considered. The continuum behaviour
can be described by a set of governing equations which are the balance
of momentum equation ($\rho \vect{a} = \Div{\tens{\sigma}} + \rho
\vect{b}$), the compatibility equation ($\dot{\tens{\varepsilon}} =
\Grad{\vect{v}}$), the constitutive equation ($\dot{\tens{\sigma}} =
\tens{D} \colon \dot{\tens{\varepsilon}}$), and the balance of mass equation
($\frac{D \rho}{D t} = \dot{\rho} + \rho \Grad{\vect{v}}$). The balance of momentum can be
formulated in weak form by introducing an arbitrary test function $\vec{\psi}$, yielding
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \Deriv{\vec{v}}{t}\ \vec{\psi}} = -
  \Integral{3}{\tens{\sigma} \colon \Grad{\vec{\psi}}} +
  \Integral{2}{\vec{t}\ \vec{\psi}} + \Integral{3}{
  \rho\ \vec{b}\ \vec{\psi}}.
\end{equation}\\

In order to arrive to a finite set of equations, the continuum domain
$\Omega$ is discretized with a finite sum of particles (material points), each one 
represent a part of the domain $\Omega_p$ with $p = 1,2\ldots ,N_p$
where $N_p$ is the number of particles. The index $p$ is
used for the particles, while the index $I$ is reserved in this
notation for denoting grid nodes. The material point $\vec{x}_p$ is
defined at the centroid of $\Omega_p$. Further, a computational grid
is formed of a finite set of grid points with coordinates $\vec{x}_I\
, I = 1,2\ldots ,N_n$, where $N_n$ is the number of grid nodes. Each material point is assigned with initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,  $\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$. The problem is then integrated using a discrete number of time
steps $k = 1\ldots ,Nt$, where $k$ is the current time step and $N_t$
is the total number of time steps.  So employing the definition of the
material integral, individual terms of \eqref{eq:BalanceMomentum_wf}
are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \vec{a}\ \vec{\psi}} =& \sum^{N_p}_{p}
   \dot{\vec{v}}_{p}\ \vec{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \Grad{\vec{\psi}}} = \sum^{N_p}_{p}
   \tens{\sigma}_{p}\ \Grad{\tens{\psi}_{p}}\ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b}\ \vec{\psi} } = \sum^{N_p}_{p}
  \vec{b}_{p}\ \vec{\psi}_{p}\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vec{\psi}} = \Integral{2}{\rho\
      \vec{t}^s\ \vec{\psi}} = \sum^{N_p}_{p} \vec{t}^s_{p}\ \vec{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
where $h$ is the thickness of the continuum in a 2D case. Introducing \eqref{eq:particle_acceleration_forces},
\eqref{eq:particle_internal_forces}, \eqref{eq:particle_body_forces},
and \eqref{eq:particle_load_forces} in \eqref{eq:BalanceMomentum_wf}
we reach to the particle balance of forces of the continuum,
\begin{equation}
  \label{eq:particle_balance_forces2}
  \sum^{N_p}_{p}\vec{\psi}_{p}\ m_p\ \dot{\vec{v}}_{p} = \sum^{N_p}_{p}
  \vec{\psi}_{p}\ m_p\ \vec{b}_{p} + \sum^{N_p}_{p}
  \vec{\psi}_{p}\ m_p\ h^{-1}\ \bar{\vec{t}}^s_{p} -
  \sum^{N_p}_{p} \Grad{\vec{\psi}_{p}}\ \frac{m_p}{\rho_p}\
  \tens{\sigma}_{p}.
\end{equation}\\

Approximating the virtual displacement of the particle $\vec{\psi}_{p}$
using a suitable approximation scheme we get \eqref{eq:VirtualDisplacementParticle}.
\begin{equation}
  \label{eq:VirtualDisplacementParticle}
  \vec{\psi}_{p} = N_{Ip}\cdot \vect{1} = N_{Ip},
\end{equation}
where $N_{Ip} = N_I(\vec{x}_p)$ it is the nodal value of the shape
function evaluated in the material point coordinates denoted by $\vec{x}_p$. Substituting \eqref{eq:VirtualDisplacementParticle} in
\eqref{eq:particle_balance_forces2} we get :
\begin{equation}
  \label{eq:particle_balance_forces3}
  \dot{\vec{p}}_{I}= \tens{m}_{IJ}\dot{\vec{v}}_{J} = \vec{f}_{I}^{int} + \vec{f}_{I}^{ext},
\end{equation}
where $\dot{\vec{p}}_{I}$ is the rate of momentum at grid node $I$, the nodal mass matrix $\tens{m}_{IJ}$ is,
\begin{equation}
  \label{eq:particle_nod_mass}
  \tens{m}_{IJ} =
  \sum^{Np}_{p} N_{Ip} m_p N_{Jp},
\end{equation}
and internal and external forces are computed as follows,
\begin{equation}
  \label{eq:nodal_internal_forces}
  \vec{f}_{I}^{int} = - \sum^{N_p}_{p} \Grad{N_{Ip}}
  \tens{\sigma}_{p} \frac{m_p}{\rho_p}
\end{equation}
\begin{equation}
  \label{eq:nodal_external_forces}
  \vec{f}_{I}^{ext} = \sum^{N_p}_{p} N_{Ip} m_p \vec{b}_{p} + \sum^{N_p}_{p}
  N_{Ip} m_p h^{-1} \vec{t}^s_{p}
\end{equation}
where $\tens{\sigma}_{p} = \tens{\sigma}_{p}(\tens{\varepsilon}_{p})$
is the particle $p$ stress field, which can be integrated employing
the suitable constitutive model. The strain tensor is updated employing the rate of stress tensor $\dot{ \tens{\varepsilon}}_{p}$ used to update the
strain tensor is as follows \eqref{eq:IncrStrainPoint}.
\begin{equation}
  \label{eq:IncrStrainPoint}
  \dot{\tens{\varepsilon}_{p}} = \frac{\Delta
    \tens{\varepsilon}_{p}}{\Delta t} =
  \frac{1}{2}\left[\Grad{N_{Ip}}\ \vec{v}_{I} + \Grad{N_{Ip}}^T\
    \vec{v}_{I}\right].
\end{equation}
Later, imposing $\frac{D \rho}{D t} = 0$, we ensures the mass
conservation and update the density field.
\begin{equation}
  \label{eq:MassConservation}
\dot{\rho} = - \rho\ \mathit{tra} \left( \dot{\tens{\varepsilon}} \right)
\end{equation}

@@

To solve equations \eqref{eq:SimplyNodalMomentum} and
\eqref{eq:Updated_Lagrangian}, a time integration scheme is
required. It has been described in detail by many researchers
\cite{Sulsky1994}, \cite{Bardenhagen2002}, \cite{thesis_Andersen_2009} and summarized
in Figure \ref{fig:MPM_algorithm}. Other authors have proposed many
others time integration alternatives like
\cite{Guilkey2003}, \cite{Tran2019e}, \cite{Charlton2017}. In the first publication on
the MPM \cite{Sulsky1994}, the nodal acceleration in employed to
update the particles as
\begin{equation}
  \label{eq:Sulsky-1994-UL-v-x}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\
  \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{a}_{I}^{k},\quad and\quad
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{v}_{I}^{k}.
\end{equation}

However, in Andersen (2009)\cite{thesis_Andersen_2009} this
algorithm has been shown to be numerically unstable due to that
$\vect{f}_I^{int,k}$ can be finite for an infinitesimal nodal mass
$\tens{m}$. This can lead to numerical issues when nodal acceleration
is obtained for evaluating \eqref{eq:Sulsky-1994-UL-v-x}. Hence, a
corrected version of this algorithm in shown in Zhang {\it et al.}
(2016)\cite{Zhang_book_2016}
\begin{equation}
  \label{eq:Zhang-2016-UL-v-x}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\
  \sum^{N_n}_{I}\frac{N_{Ip}^{k}\ \vec{f}_{I}^{k}}{\tens{m}_I},\quad and\quad
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ \sum^{N_n}_{I} \frac{N_{Ip}^{k}\ \vec{p}_{I}^{k}}{\tens{m}_I}.  
\end{equation}

Later Tran \& Solowski (2019)\cite{Tran2019e} presented a
generalized-$\alpha$ scheme for the MPM inspired in the explicit time
integration algorithm proposed by Chung \& Hulbert
(1993)\cite{Geranlized_alpha_1993}, but with the particularity that
the acceleration is evaluated both in the beginning and the end of the
time step.
\begin{align}
  \label{eq:Tran-2019-GA-v}
  & \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\
    \sum^{N_n}_{I} N_{Ip}^{k}\ \left[(1 - \gamma)\ \vect{a}_I^{k} +
    \gamma\ \vect{a}_I^{k+1} \right]\\
  \label{eq:Tran-2019-GA-x}
  & \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\
    \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{v}_{I}^{k} + \Delta t^2
    \sum^{N_n}_{I}N_{Ip}^{k} \left[ (\frac{1}{2} - \beta)\
    \vec{a}_{I}^{k} + \beta\ \vec{a}_{I}^{k+1} \right]
\end{align}
and
\begin{equation}
  \label{eq:Tran-2019-GA-a}
   \vect{a}_p^{k+1} = \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{a}_{I}^{k+1}.
\end{equation}

This scheme has prof to damps out the higher frequency noises
\cite{Tran2019e}. But it can present the same numerical instabilities
as in \eqref{eq:Sulsky-1994-UL-v-x} when nodal masses become infinitesimal.
Here we propose a explicit predictor-corrector scheme for MPM based on the Newmark a-form with
$\gamma = 0.5$ and $\beta = 0$ which is the central difference
explicit. This method is devoted to solve a system of equations of type
\begin{equation*}
  \Matrix{M}_{IJ}\ddot{\Vector{d}}_{J} + \Matrix{C}_{IJ}\dot{\Vector{d}}_{J} +
  \Matrix{K}_{IJ}\Vector{d}_{J} = \Vector{F}_{I}
\end{equation*}
Nevertheless it can be applied for an isolated mass, therefore it is
possible to apply this methods successfully in the MPM framework as
was proved by \cite{Tran2019e}. Taking the predictor
definition from the classic literature \cite{Hughes2000} and
calculating particles velocity and acceleration as described in 
\cite{Zhang_book_2016}, we get the following expression for the
\textbf{predictor} stage for both velocity and displacement.
\begin{align}
  \label{eq:Predictor-displacement}
  &\vec{x}_p^{n+1} = \vec{x}_p^n + \Delta t\ \underbrace{\sum^{N_n}_{I}\frac{\vec{p}_{I}^{k}\ N_{Ip}^k}{\tens{m}_I^k}}_{\vec{v}_p^{n}} + \frac{1}{2}\Delta t^2\ \underbrace{\sum^{N_n}_{I}\frac{\vec{f}_{I}^{k}\ N_{Ip}^k}{\tens{m}_I^k}}_{\vec{a}_p^{n}}\\
  \label{eq:Predictor-velocity}
  &\vec{v}_p^{n+1} = \vec{v}_p^n + (1 - \gamma)\ \Delta t\ \underbrace{\sum^{N_n}_{I}\frac{\vec{f}_{I}^{k}\ N_{Ip}^k}{\tens{m}_I^k}}_{\vec{a}_p^{n}}        
\end{align}

Consequently, the \textbf{corrector} stage for the particles velocity is in the
following way
\begin{equation*}
  \vec{v}_{p}^{k+1} = \vec{v}_{p}^{k+1} +
  \gamma\ \Delta t\ \underbrace{\sum^{N_n}_{I}\frac{\vec{f}_{I}^{k+1}\ N_{Ip}^{k+1}}{\tens{m}_I^{k+1}}}_{\vec{a}_p^{n+1}}  
\end{equation*}

@@

>>>>

\message{ !name(main.tex)}%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
% flush right qed marks, e.g. at end of proof
\smartqed

@@

% Divergence
\newcommand{\Div}[1]{
  \ensuremath{\nabla \cdot {#1}}
}
% Gradient
\newcommand\Grad[1]{\nabla{#1}}
\newcommand\GradS[1]{\nabla^s{#1}}
\newcommand\GradT[1]{\nabla^T{#1}}

@@

% Integral
\newcommand{\Integral}[2]{
  \IfStrEqCase{#1}{
    {2}{\ensuremath{\int_{\varGamma_d}{#2}\ d\varGamma}}
    {3}{\ensuremath{\int_{\varOmega}{#2}\ d\varOmega}}
  }
}

@@

\message{ !name(main.tex) !offset(-3) }

@@

\section{Notation}
\label{sec:Notation}
In what follows, we will adopt the following convention:
\begin{table}[h]
  \centering
  \caption{Physical variables involved in the problem}
  \label{tab:notation-table}
  \begin{tabular}{c c}
    $\rho$ & Density field \\
    $\vect{a}$ & Acceleration field \\
    $\vect{v}$ & Velocity field \\
    $\vect{u}$ & Displacement field \\
    $\tens{\sigma}$ & Cauchy stress tensor \\
    $\tens{\varepsilon}$ & Cauchy strain tensor\\
    $\tens{D}$ & Constitutive tensor \\
  \end{tabular}
\end{table}

The subscript $p$ is used to define a particle variable. While the subscript $I$ is reserved in this
notation for denoting nodal variables.

\section{Derivation of the MPM}
\label{sec:derivation-mpm}
In this section, we provide an overview of the standard explicit MPM
algorithm as presented by \cite{Sulsky1994}. The method has three main
steps: a variational recovery process to project particle data to the
nodes of the grid, a Eulerian step where balance of momentum is
compute by solving internal and external forces in the nodes, and
finally the advection of the Lagrangian particles. We will first give
a introduction to the governing equations and then follow by an
algorithmic description of the process.

\subsection{Governing equations }
\label{sec:governing-equations}
In the MPM approach a continuum is considered. The continuum behaviour
can be described by a set of governing equations which are

the balance of momentum equation
\begin{equation}
  \label{eq:Balance-momentum}
\rho \vect{a} = \Div{\tens{\sigma}} + \rho \vect{b},
\end{equation}
where $\Div{\qed}$ denotes the divergence operator. To ensure that the
velocity field is compatible with the strain field is necessary that
them satisfy the compatibility equation,
\begin{equation}
  \label{eq:Compatibility-equation}
  \dot{\tens{\varepsilon}} = \GradS{\vect{v}} = \frac{1}{2}\left(
    \Grad{\vect{v}} +  \GradT{\vect{v}} \right),
\end{equation}
where $\GradS{\qed}$ denotes the symmetric part of the gradient
operation. Thorough the constitutive equation, the rate of the Cauchy stress
tensor is defined to be
\begin{equation}
  \label{eq:Constitutive-equation}
\dot{\tens{\sigma}} = \tens{D} \colon \dot{\tens{\varepsilon}},
\end{equation}
where $\tens{D}$ is the classical elasticity tensor. Finally we
introduce the total derivative expression over the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
\frac{D \rho}{D t} = \dot{\rho} + \rho \Grad{\vect{v}}.
\end{equation}
Where if we impose $\frac{D \rho}{D t}$ to be zero, the conservation
of mass is guarantee.

\subsection{Variational formulation}
\label{sec:variational-formulation}
To write the variational statement of the problem, let us define a
virtual velocity field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vect{v}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vect{v}^{\psi} \in
  \mathcal{H}^1 \mid \vect{v}^{\psi} = \vect{0}\ \text{on}\ \Gamma_d \}.
\end{equation}
And which satisfies that the Cauchy sequences are convergent in $\varOmega$
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vect{v}^{\psi}} < \infty\ \quad\text{and}\quad \Integral{3}{\Grad{\vect{v}^{\psi}}} < \infty
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary value problem of elasticity is the function $\vect{v} \in
\mathcal{H}^1_0$ such that, for $\vect{v}^{\psi} \in
\mathcal{H}^1_0$, the following holds:
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \Deriv{\vec{v}}{t}\ - \vec{b} \right) \cdot \vec{v}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{v}^{\psi}} - \Integral{3}{\tens{\sigma} \cdot
    \tens{\varepsilon}^{\psi}},
\end{equation}\\
therefore (\ref{eq:BalanceMomentum_wf}) is the balance of momentum formulated in
its weak form.

\subsection{Galerkin procedure}
\label{sec:Galerkin-procedure}
In order to arrive to a finite set of equations, the continuum domain
$\Omega$ is discretized with a finite sum of material points (in the
following particles), each one represent a part of the discretized domain $\Omega_p$ with $p = 1,2\ldots ,N_p$
where $N_p$ is the number of particles. The material point $\vec{x}_p$ is
defined at the centroid of each $\Omega_p$. Further, a computational grid
is formed of a finite set of grid points with coordinates $\vec{x}_I\
, I = 1,2\ldots ,N_n$, where $N_n$ is the number of grid nodes.

\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Discretization_MPM}
  }
  \caption{MPM discretization.}
  \label{fig:MPM-discretization}
\end{figure}

Each material point is assigned with initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,  $\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$. The problem is then integrated using a discrete number of time
steps $k = 1\ldots ,Nt$, where $k$ is the current time step and $N_t$
is the total number of time steps.  So employing the definition of the
material integral, individual terms of \eqref{eq:BalanceMomentum_wf}
are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \vec{a}\ \vect{v}^{\psi}} =& \sum^{N_p}_{p}
   \dot{\vec{v}}_{p}\ \vect{v}^{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \Grad{\vec{\psi}}} = \sum^{N_p}_{p}
   \tens{\sigma}_{p}\ \Grad{\tens{\psi}_{p}}\ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b}\ \vect{v}^{\psi} } = \sum^{N_p}_{p}
  \vec{b}_{p}\ \vect{v}^{\psi}_p\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vect{v}^{\psi}} = \Integral{2}{\rho\
      \vec{t}^s\ \vect{v}^{\psi}} = \sum^{N_p}_{p} \vec{t}^s_{p}\ \vect{v}^{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
where $h$ is the thickness of the continuum in a 2D case. Introducing \eqref{eq:particle_acceleration_forces},
\eqref{eq:particle_internal_forces}, \eqref{eq:particle_body_forces},
and \eqref{eq:particle_load_forces} in \eqref{eq:BalanceMomentum_wf},
and approximating the virtual displacement of the particle $p$ as
$\vec{\psi}_{p} = N_I(\vec{x}_p)\cdot \vect{1} = N_{Ip}$. We reach to
the particle balance of forces of the continuum,
\begin{equation}
  \label{eq:particle_balance_forces3}
  \dot{\vec{p}}_{I}= \tens{m}_{IJ}\dot{\vec{v}}_{J} = \vec{f}_{I}^{int} + \vec{f}_{I}^{ext},
\end{equation}
where $\dot{\vec{p}}_{I}$ is the rate of momentum at grid node $I$, the nodal mass matrix $\tens{m}_{IJ}$ is,
\begin{equation}
  \label{eq:particle_nod_mass}
  \tens{m}_{IJ} =
  \sum^{Np}_{p} N_{Ip} m_p N_{Jp},
\end{equation}
and internal and external forces are computed as follows,
\begin{equation}
  \label{eq:nodal_internal_forces}
  \vec{f}_{I}^{int} = - \sum^{N_p}_{p} \Grad{N_{Ip}}
  \tens{\sigma}_{p} \frac{m_p}{\rho_p}
\end{equation}
\begin{equation}
  \label{eq:nodal_external_forces}
  \vec{f}_{I}^{ext} = \sum^{N_p}_{p} N_{Ip} m_p \vec{b}_{p} + \sum^{N_p}_{p}
  N_{Ip} m_p h^{-1} \vec{t}^s_{p}
\end{equation}
where $\tens{\sigma}_{p} = \tens{\sigma}_{p}(\tens{\varepsilon}_{p})$
is the particle $p$ stress field, which can be integrated employing
the suitable constitutive model. The strain tensor is updated employing the rate of stress tensor $\dot{ \tens{\varepsilon}}_{p}$ used to update the
strain tensor is as follows \eqref{eq:IncrStrainPoint}.
\begin{equation}
  \label{eq:IncrStrainPoint}
  \dot{\tens{\varepsilon}_{p}} = \frac{\Delta
    \tens{\varepsilon}_{p}}{\Delta t} =
  \frac{1}{2}\left[\Grad{N_{Ip}}\ \vec{v}_{I} + \Grad{N_{Ip}}^T\
    \vec{v}_{I}\right].
\end{equation}
Later, imposing $\frac{D \rho}{D t} = 0$, we ensures the mass
conservation and update the density field.
\begin{equation}
  \label{eq:MassConservation}
\dot{\rho} = - \rho\ \mathit{tra} \left( \dot{\tens{\varepsilon}} \right)
\end{equation}

@@

To solve equations \eqref{eq:SimplyNodalMomentum} and
\eqref{eq:Updated_Lagrangian}, a time integration scheme is
required. It has been described in detail by many researchers
\cite{Sulsky1994}, \cite{Bardenhagen2002}, \cite{thesis_Andersen_2009} and summarized
in Figure \ref{fig:MPM_algorithm}. Other authors have proposed many
others time integration alternatives like
\cite{Guilkey2003}, \cite{Tran2019e}, \cite{Charlton2017}. In the first publication on
the MPM \cite{Sulsky1994}, the nodal acceleration in employed to
update the particles as
\begin{equation}
  \label{eq:Sulsky-1994-UL-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\
  \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{a}_{I}^{k}
\end{equation}
\begin{equation}
  \label{eq:Sulsky-1994-UL-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{v}_{I}^{k}.
\end{equation}
However, in Andersen (2009)\cite{thesis_Andersen_2009} this
algorithm has been shown to be numerically unstable due to that
$\vect{f}_I^{int,k}$ can be finite for an infinitesimal nodal mass
$\tens{m}$. This can lead to numerical issues when nodal acceleration
is obtained for evaluating \eqref{eq:Sulsky-1994-UL-x},\eqref{eq:Sulsky-1994-UL-v}. Hence, a
corrected version of this algorithm in shown in Zhang {\it et al.}
(2016)\cite{Zhang_book_2016}
\begin{equation}
  \label{eq:Zhang-2016-UL-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ \sum^{N_n}_{I} \frac{N_{Ip}^{k}\ \vec{p}_{I}^{k}}{\tens{m}_I}.  
\end{equation}
\begin{equation}
  \label{eq:Zhang-2016-UL-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\
  \sum^{N_n}_{I}\frac{N_{Ip}^{k}\ \vec{f}_{I}^{k}}{\tens{m}_I},
\end{equation}
Later Tran \& Solowski (2019)\cite{Tran2019e} presented a
generalized-$\alpha$ scheme for the MPM inspired in the explicit time
integration algorithm proposed by Chung \& Hulbert
(1993)\cite{Geranlized_alpha_1993}, but with the particularity that
the acceleration is evaluated both in the beginning and the end of the
time step.
\begin{equation}
  \label{eq:Tran-2019-GA-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\
  \sum^{N_n}_{I} N_{Ip}^{k}\ \left[(1 - \gamma)\ \vect{a}_I^{k} +
    \gamma\ \vect{a}_I^{k+1} \right],\\
\end{equation}
\begin{equation}
\label{eq:Tran-2019-GA-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + 
  \sum^{N_n}_{I}N_{Ip}^{k} \left[ \Delta t\ \vec{v}_{I}^{k}+ \Delta t^2\left( (\frac{1}{2} - \beta)\
    \vec{a}_{I}^{k} + \beta\ \vec{a}_{I}^{k+1} \right) \right]
\end{equation}
\begin{equation}
  \label{eq:Tran-2019-GA-a}
  \vect{a}_p^{k+1} = \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{a}_{I}^{k+1}.
\end{equation}

This scheme has prof to damps out the higher frequency noises
\cite{Tran2019e}. But it can present the same numerical instabilities
as in \eqref{eq:Sulsky-1994-UL-x},\eqref{eq:Sulsky-1994-UL-v} when nodal masses become infinitesimal.
Here we propose a explicit predictor-corrector scheme for MPM based on the Newmark a-form with
$\gamma = 0.5$ and $\beta = 0$ which is the central difference
explicit. This method is devoted to solve a system of equations of type
\begin{equation*}
  \Matrix{M}_{IJ}\ddot{\Vector{d}}_{J} + \Matrix{C}_{IJ}\dot{\Vector{d}}_{J} +
  \Matrix{K}_{IJ}\Vector{d}_{J} = \Vector{F}_{I}
\end{equation*}
Nevertheless it can be applied for an isolated mass, therefore it is
possible to apply this methods successfully in the MPM framework as
was proved by \cite{Tran2019e}. Taking the predictor
definition from the classic literature \cite{Hughes2000} and
calculating particles velocity and acceleration as described in 
\cite{Zhang_book_2016}, we get the following expression for the
\textbf{predictor} stage for both velocity and displacement.
\begin{equation}
  \label{eq:5}
  \vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
  \underbrace{\sum^{N_n}_{I}\frac{\vec{p}_{I}^{k}\
      N_{Ip}^k}{\tens{m}_I^k}}_{\vec{v}_p^{k}} + \frac{1}{2}\Delta
  t^2\ \underbrace{\sum^{N_n}_{I}\frac{\vec{f}_{I}^{k}\
      N_{Ip}^k}{\tens{m}_I^k}}_{\vec{a}_p^{k}}
\end{equation}
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{v}_p^{k+1} = \vec{v}_p^n + (1 - \gamma)\ \Delta t\
  \underbrace{\sum^{N_n}_{I}\frac{\vec{f}_{I}^{k}\
      N_{Ip}^k}{\tens{m}_I^k}}_{\vec{a}_p^{k}}
\end{equation}

Consequently, the \textbf{corrector} stage for the particles velocity is in the
following way
\begin{equation}
  \vec{v}_{p}^{k+1} = \vec{v}_{p}^{k+1} +
  \gamma\ \Delta t\ \underbrace{\sum^{N_n}_{I}\frac{\vec{f}_{I}^{k+1}\ N_{Ip}^{k+1}}{\tens{m}_I^{k+1}}}_{\vec{a}_p^{k+1}}  
\end{equation}

@@

\message{ !name(main.tex) !offset(-1152) }
